\chapter* {Acknowledgments}
\addcontentsline{toc}{section}{\numberline{}\hspace{-.35in}{\bf
Acknowledgments}}

It has been my utmost privilege to be given this opportunity to pursue a Ph.D degree. First, I would like to express my sincere gratitude to my supervisor, Prof. A. S. Madhukumar for his invaluable mentorship throughout my Ph.D candidature. The countless discussions that I have had with him have played a considerable role in orientating my research direction in the vast area of wireless communications. As a fair and honest critic of my work, Prof. Madhu has enabled me to have a better understanding of my research area through his insights and encouragement. His expectation of high standard and rigour has also been instrumental in raising the quality of my work.

Next, I would like to thank my co-supervisor, Dr Anoop Kumar Krishna, for his support of my Ph.D project. With his long experience and deep expertise in the industry, Dr Anoop's comments and suggestions have been crucial in improving the practical relevance of my work. I am also grateful for Dr Anoop's help in enabling me to have an enriching internship at Airbus Singapore Pte Ltd before the start of my candidature. The internship allowed me to experience upstream research and development, and also served as an introduction to my Ph.D topic.

I would also like to thank Dr Rajendra Prasad Sirigina for his significant role in imparting valuable knowledge to me, particularly in the first two years of my candidature. It was only through Dr Rajendra's guidance that I was able to grasp key mathematical concepts and techniques that were necessary for my work. I am also particularly grateful to Dr Rajendra for the advice and encouragement dispensed in the course of the review process for all my papers, and the help in critiquing my work, despite his busy schedule.

I want to extend my thanks to Airbus Singapore Pte Ltd and the Singapore Economic Development Board (EDB) for the support and funding of this Ph.D project, and also to Nanyang Technological University for the conducive environment in the School of Computer Science and Engineering. 

Last, but not least, I would like to thank my parents and Li Ying for being my pillar of support throughout my Ph.D candidature. Their love, care, and encouragement have been the key to sustaining my mental strength throughout my intellectual pursuit these past four years. 



