\chapter* {List of Publications}
%\addcontentsline{toc}{chapter}{\numberline{}\hspace{-0.24in}{\bf
%Publication}}

\section* {International Conferences}
\begin{enumerate}
		\item Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Mitigating Cellular Interference in Uplink UAV Communications,” Submitted to IEEE 92nd Veh. Technol. Conf. (VTC Fall) 2020 for possible publication.
    \item Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Impact of Cellular Interference on Uplink UAV Communications,” in Proc. IEEE 91st Veh. Technol. Conf. (VTC Spring), Antwerp, Belgium, 2020
		\item	Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Capacity Characterization of Uplink NOMA in Multi-UAV Networks,” in Proc. IEEE 91st Veh. Technol. Conf. (VTC Spring), Antwerp, Belgium, 2020.
		\item	Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Downlink NOMA in Multi-UAV Networks over Bivariate Rician Shadowed Fading Channels,” in Proc. IEEE 90th Veh. Technol. Conf. (VTC Fall), Hawaii, USA, 2019.
		\item	Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “An Outage Probability Analysis of Full-Duplex NOMA in UAV Communications,” in Proc. IEEE Wireless Commun. Netw. Conf. (WCNC), Marrakech, Morocco, 2019, pp. 1–5.
		\item Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Hybrid-Duplex Systems for UAV Communications Under Rician Shadowed Fading,” in Proc. IEEE 88th Veh. Technol. Conf. (VTC Fall), Chicago, USA, 2018.
		\item	Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Hybrid-Duplex based Control and Non-Payload Communication Systems for UAVs: An Outage Analysis,” in Proc. IEEE/AIAA 37th Digit. Avionics Syst. Conf. (DASC), London, UK, 2018.
		\item	Tan Zheng Hui Ernest, Rajendra Prasad Sirigina, Anoop Kumar Krishna, and A. S. Madhukumar, “On the Performance Analysis of Hybrid-Duplex Systems for Aeronautical Communications,” in Proc. IEEE 87th Veh. Technol. Conf. (VTC Spring), Porto, Portugal, 2018.
		\item Tan Zheng Hui Ernest, Anoop Kumar Krishna, A. S. Madhukumar, and Rajendra Prasad Sirigina, “On the Efficiency Improvements to Aeronautical Waveforms and Integrated Modular Avionics Systems,” in Proc. IEEE/AIAA 35th Digit. Avionics Syst. Conf. (DASC), 2016. IEEE, Sacramento, CA, USA, 2016, pp. 1–8.
\end{enumerate}

\section* {International Journals}
\begin{enumerate}
		\item	Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “NOMA-Aided Multi-UAV Communications in Full-Duplex Heterogeneous Networks,” Submitted to IEEE Systems Journal for possible publication.
		\item Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “NOMA-aided UAV Communications over Correlated Rician Shadowed Fading Channels,” Submitted to IEEE Transactions on Signal Processing for possible publication.
		\item	Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Addressing Spectrum Efficiency Through Hybrid-Duplex UAV Communications: Challenges and Opportunities,” Elsevier Vehicular Communications, vol. 24, August 2020.
		\item Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Hybrid-Duplex Communications for Multi-UAV Networks: An Outage Probability Analysis,” IEEE Communications Letters, vol. 23, no. 10, pp. 1831–1835, October 2019.
		\item Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “A Power Series Approach for Hybrid-Duplex UAV Communication Systems under Rician Shadowed Fading,” IEEE Access, vol. 7, pp. 76949–76966, June 2019.
		\item Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “Outage Analysis and Finite SNR Diversity-Multiplexing Tradeoff of Hybrid-Duplex Systems for Aeronautical Communications,” IEEE Transactions on Wireless Communications, vol. 18, no. 4, pp. 2299–2313, April 2019.
		\item	Tan Zheng Hui Ernest, A. S. Madhukumar, Rajendra Prasad Sirigina, and Anoop Kumar Krishna, “A Hybrid-Duplex System with Joint Detection for Interference-Limited UAV Communications,” IEEE Transactions on Vehicular Technology, vol. 68, no. 1, pp. 335–348, January 2019.
\end{enumerate}
